---
title: "District 7"
author: "Alex Reich"
date: "2024-04-24"
output: html_document
---
Aside:
Argh, I had to upgrade to R 4.4.0 to get lme4 (the matrix package, specifically) to work. But now the r markdown settings are different and I can't default to the home directory being the project. So I will set that here:
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/agreich/OneDrive - State of Alaska/Desktop/Shrimp/Shrimp tickets (pot shrimp)/SE_pot_shrimp_CPUE')
```



Anyway, let's get started. This document will look at management unit District 7 for pot (spot) shrimp standardization. It will load in the (pre-wrangled) data, wrangle further for district 1, and then some EDA (focusing on variables of interest and correlation). The analysis will be by district with an HGAM (hierarchical GAM) with the variables of interest as fixed effects and Analysis Area as a random effect. I will do model selection on a few versions of this HGAM. I will then make graphs of the standardized CPUE (by district and) and whatever else Max wants. THEN, I will run individual models by analysis areas, ideally with an automated function, and compare the graphs of those automated functions to the cpue of the ranef model.This will help me decide if models should be by MANAGEMENT UNIT (larger) or by ANALYSIS AREA (smaller). In hindsight, I should start with District 7, since I have explored the Upper Ernest sound data extensively.Ok, now the RMD is named accordingly 


Load libraries
```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(mgcv)
```


Load the data and select for D7
```{r}
wrangled_shrimp <- read.csv("Data/wrangled shrimp focus years.csv")
D7 <- wrangled_shrimp %>% filter(Management_unit == "District 7") %>%
  mutate(Analysis.Area = factor(Analysis.Area),
         ADFG.Number = factor(ADFG.Number),
         Season.Ref = factor(Season.Ref)
         )#make things factors for the gams

unique(D7$Analysis.Area)

```


District 7 contains analysis areas:
  Bradfield Canal
  Lower Ernest Sound
  Upper Ernest Sound
  Zimovia Strait
  
  
  
## Exploratory Data Analysis for D7
Variables of interest:
Random effect: Analysis.Area
Fixed effects: jdate (smoothed and... cyclical?), vessel # (the count of vessels), ADFG # (individual vessel (fixed or random effect??)), Season aka year, 
--If I do not include jdate, simplifies things?
Response: nominal cpue

Q: should I do something besides a log-transformation?
THIN OUT
```{r}

#cpue dist
ggplot(D7) + aes(x=CPUE_nom) + geom_density()
ggplot(D7) + aes(x=log(CPUE_nom+0.001)) + geom_density()
qqnorm(log(D7$CPUE_nom)) #that kind of looks bad. Maybe somehting other than log-trans?
qqnorm(D7$CPUE_nom)

ggplot(D7) + aes(x=CPUE_nom) + geom_density()+facet_wrap(~Analysis.Area)
ggplot(D7) + aes(x=log(CPUE_nom+0.001)) + geom_density()+facet_wrap(~Analysis.Area)

#cpue by year
ggplot(D7) + aes(x=factor(Season.Ref), y=CPUE_nom) + geom_point() #ther is a big outlier
ggplot(D7) + aes(x=factor(Season.Ref), y=CPUE_nom) + geom_boxplot(outliers = F)

ggplot(D7) + aes(x=factor(Season.Ref), y=CPUE_nom) + geom_boxplot(outliers = F) + facet_wrap(~Analysis.Area) #cpue varies... cyclically with time. Is season a fixed effect? I want to estimate it, so probs not. its def not linear.

#catch by year
ggplot(D7) + aes(x=factor(Season.Ref), y=total_weight) + geom_boxplot() #this is total weight per fish ticket.
ggplot(D7) + aes(x=factor(Season.Ref), y=max_pots_2) + geom_boxplot()
ggplot(D7) + aes(x=factor(Season.Ref), y=max_pots_2) + geom_boxplot(outliers=F)

ggplot(D7) + aes(x=factor(Season.Ref), y=total_weight) + geom_boxplot() + facet_wrap(~Analysis.Area) #max pots per fish ticket
ggplot(D7) + aes(x=factor(Season.Ref), y=max_pots_2) + geom_boxplot(outliers=F) + facet_wrap(~Analysis.Area)


#cpue and vessel count
ggplot(D7) + aes(x=factor(vessel_count_mgmt_u), y=log(CPUE_nom)) + geom_boxplot()
ggplot(D7) + aes(x=factor(vessel_count_mgmt_u), y=log(CPUE_nom)) + geom_boxplot(outliers=F)
ggplot(D7) + aes(x=factor(vessel_count_mgmt_u), y=log(CPUE_nom)) + geom_boxplot(outliers=F) + facet_wrap(~Analysis.Area)

#cpue and district


#cpue and jdate



#correlation plot
library(corrplot)
D7_interest_cor <- D7 %>% select(CPUE_nom, jdate, vessel_count_mgmt_u)
cor_prep <- cor(D7_interest_cor)
corrplot(cor_prep)
?pairs


#check model residuals, see if we need ar1 correlation in the model
```


Drop the outlier, perhaps
```{r}
D7 <- D7 %>% filter(CPUE_nom < 100) #gonna assume that is an outlier
```



Model selection 5/15/24
```{r}

#ok can breifly test ranefs
M_ranef <-  lmer(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + Season.Ref:Analysis.Area + Analysis.Area + (1|ADFG.Number), data=D7, REML=T)


M_noranef <-  lm(log(CPUE_nom+0.001) ~ Season.Ref +vessel_count_aa + ADFG.Number + Season.Ref:Analysis.Area + Analysis.Area, data=D7)
AIC(M_ranef, M_noranef) #no ranefs

summary(M_ranef)



#global with ranef and interactions

M2 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + ADFG.Number +Analysis.Area + Season.Ref:Analysis.Area, data=D7)

M3 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + ADFG.Number + Analysis.Area + Season.Ref:Analysis.Area, data=D7)

M4 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + ADFG.Number +Analysis.Area, data=D7)

M5 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + Analysis.Area + Season.Ref:Analysis.Area, data=D7)

M_null <- lm(log(CPUE_nom+0.001) ~ Season.Ref+ Analysis.Area, data=D7)

AIC(M2, M3, M4, M5, M_null) #M3 wins
#says that vessel count does not matter...


summary(M3)

```

Impute values, and get values in log-space
Uncertainty: the method I chose for this.
```{r}
D7$logCPUE <- log(D7$CPUE_nom + 0.001)

#impute values so we dont have missing year:stat area problems
library(mice)



##################################################################################################
#lets try imputing data for area:year wihtout imputing for every single damn boat
temp2 <- expand.grid(Season.Ref = unique(D7$Season.Ref),
                       Analysis.Area = unique(D7$Analysis.Area)
                      )
D7_sub <- D7 %>% select(Season.Ref, Analysis.Area, ADFG.Number, logCPUE)
D7_sub2 <- full_join(D7_sub, temp2)
#make missing values for ADFG.Number the mode
which.max(table(D7_sub2$ADFG.Number)) #52131

mode_boat <- which.max(table(D7$ADFG.Number))

mode_val <- names(mode_boat)

D7_sub3 <- D7_sub2 %>%
  #mutate(ADFG.Number = ifelse(is.na(ADFG.Number), mode_val, ADFG.Number)) #replacing NA with the mode fishing vessel
  mutate(ADFG.Number = replace_na(ADFG.Number, mode_val)) #replace_na(list(ADFG.Number = mode_val)))

#now impute
md.pattern(D7_sub3)
imp <- mice(data = D7_sub3, m=1, maxit=1 ) #Ppm method
D7_imputated <- complete(imp)
View(complete(mice(data = D7_sub3, m=1, maxit=1, method="mean" )))
imp2 <- mice(data = D7_sub3, m=5, seed=444)

#following the vignette, experimenting
imp2$predictorMatrix #the...predictor matrix
plot(imp2) #neat. why do I not have an SD? probs because I just imputed one value in this case?
imp2$method #says the method of imputation for the only column I imputed (logCPUE)

#try using more iterations, just because(5 should be fine tho)
imp40 <- mice.mids(imp2, maxit=35, print=F)
plot(imp40) #neat

#stripplot
stripplot(imp2)
fit_imp <- with(imp2, lm()) #um.... can do my cpue standardization right here I guess #https://www.gerkovink.com/miceVignettes/Convergence_pooling/Convergence_and_pooling.html
imp2_pool <- pool(fit_imp)
#I dont think I will do the above step. I think I will use complete(imp2) to get my dataset, and roll to model selection from there.

#consider adding vessel # to the impued value? no...?
```
5/28/24: weights on areas (the size of the analysis areas when modeling!! see if you can remove the bias of additional data)
Model selection after imputation

Predict 5/15/24
```{r}
table(D7$ADFG.Number)


std_dat<- expand.grid(Season.Ref = unique(D7$Season.Ref), #do I need to extrapolate for areas/years taht do not exist??
                              Analysis.Area = unique(D7$Analysis.Area), #does not like east behm canal
                              #Hooks = round(mean(cpue_dat$Hooks),0), 
                              #Line_no = 2, #mean(cpue_dat$Line_no) unique(cpue_dat$Line_no)
                              ADFG.Number = factor(52131) #52131 appears 399 times.
                      )

pred_cpue <- predict(M3, std_dat, type = "response", se = TRUE) #does have the doubtful cases error, with both 71619  and 33175
#oh so it works for M13 but not for M12?


#OR, what if I just skipped the years:area combinations taht dont exist? Do I need to predict for those?!!!!!!!!!!!
##cant I just predict for the situations where CPUE DOES exist??

```


Standardized (predicted)  CPUE
(from internet: Note I predict the log version as exp(predict(m) + sigma^2/2) because $E exp(X) = exp(mu + sigma^2)$ where $X \sim N(0, sigma^2)$. (This changes little here though.))
See: https://stats.stackexchange.com/questions/115571/back-transforming-regression-results-when-modeling-logy
```{r}
?predict.merMod
pred<- predict(m_glob6_IR, type = "response", se = TRUE) #newdata needs something?? right?

#newdata
newdata <- expand.grid(Season.Ref = unique(D7$Season.Ref),
                               vessel_count_aa = round(mean(D7$vessel_count_aa),0),   #  unique(corr_spot_limited$jdate)... might need to expand and average this over all dates too
                               ADFG.Number = unique(D7$ADFG.Number), #table(corr_spot$ADFG.Number) #52131 #56332 #41228
                              Analysis.Area = unique(D7$Analysis.Area)
                               
 )

pred1 <- predict(m_glob6_IR, newdata, type = "response", se = TRUE)

pred_df <- data.frame(newdata, pred1$fit, pred1$se.fit)

#is this misleading? I have all vessels fising in all locations for this pred.
unique(D7 %>% filter(Analysis.Area=="Bradfield Canal") %>% select(ADFG.Number)) #30
unique(D7 %>% filter(Analysis.Area=="Lower Ernest Sound") %>% select(ADFG.Number)) #39
unique(D7 %>% filter(Analysis.Area=="Upper Ernest Sound") %>% select(ADFG.Number)) #51
unique(D7 %>% filter(Analysis.Area=="Zimovia Strait") %>% select(ADFG.Number)) #30
unique(D7$ADFG.Number) #88
###so, this might not be the best predict. Maybe I should leave the vessel part out, or choose a vessel in the middle?
###also, filter out unknown vessel at some point please

#what is, like, a very average fishing vessel, that is in all locations?
#or I could use jsut the fishing vessels that overlap? WHAT DID PHIL AND TYLER DO?!?!?!?!
###oh. see Spot_shrimp_ernest_clean_4.R code for an option.starts at 894. I weigh the ADFG.Number based on how many timest it appears in the dataset. How does that work for multiple areas tho??


#newdata
newdata2 <- expand.grid(Season.Ref = unique(D7$Season.Ref),
                               vessel_count_aa = round(mean(D7$vessel_count_aa),0),   #  unique(corr_spot_limited$jdate)... might need to expand and average this over all dates too
                              # ADFG.Number = unique(D7$ADFG.Number), #table(corr_spot$ADFG.Number) #52131 #56332 #41228
                              Analysis.Area = unique(D7$Analysis.Area)
                               
 )

pred2 <- predict(m_glob6_IR, newdata2, type = "response", se = TRUE) #ok so adfg numver should be something


#also, remember that exp() will give you the mean, so use the adjustment to get the median plz!!!!
```




Analysis and model testing: individual analysis areas
```{r}
#Bradfield canal

#Lower ernest sound

#upper ernest sound

#Zimovia strait


```

Comparison graphs
```{r}

```

Results graphs
```{r}

```

Aside:
IS ADFG vessel accounted for?
-in email max says:

This is where shrimp survey data is located.
https://www.adfg.alaska.gov/analytics/saw.dll?Answers&path=%2Fshared%2FCommercial%20Fisheries%2FRegion%20I%2FInvertebrates%2FSurvey%2FShrimp%2FSpecimens%20with%20Pot%20Data

Survey catches are also reported in fish ticket data, recorded as Fishery: Test Fish, and Permit Holder: ADFG Sitka. These survey catches should be removed from calculations in CPUE and total reportable harvest.- I DO NOT SEE THESE REPORTED THO


_______

Take away from KAtie meeting:
focus on shrimp ticket for now.
Ask Max at what level does he make decisions. When he closes an area, what is that based off of?

