---
title: "District 7"
author: "Alex Reich"
date: "2024-04-24"
output: html_document
---
I do not  remember if I need to use the following code:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Anyway, let's get started. This document will look at management unit District 7 for pot (spot) shrimp standardization. It will load in the (pre-wrangled) data, wrangle further for district 1, and then some EDA (focusing on variables of interest and correlation). The analysis will be by district with an HGAM (hierarchical GAM) with the variables of interest as fixed effects and Analysis Area as a random effect. I will do model selection on a few versions of this HGAM. I will then make graphs of the standardized CPUE (by district and) and whatever else Max wants. THEN, I will run individual models by analysis areas, ideally with an automated function, and compare the graphs of those automated functions to the cpue of the ranef model.This will help me decide if models should be by MANAGEMENT UNIT (larger) or by ANALYSIS AREA (smaller). In hindsight, I should start with District 7, since I have explored the Upper Ernest sound data extensively.Ok, now the RMD is named accordingly 


Load libraries
```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(mgcv)
```


Load the data and select for D7
```{r}
wrangled_shrimp <- read.csv("Data/wrangled shrimp focus years.csv")
D7 <- wrangled_shrimp %>% filter(Management_unit == "District 7")

unique(D7$Analysis.Area)
```


District 7 contains analysis areas:
  Bradfield Canal
  Lower Ernest Sound
  Upper Ernest Sound
  Zimovia Strait
  
  
  
## Exploratory Data Analysis for D7
Variables of interest:
Random effect: Analysis.Area
Fixed effects: jdate (smoothed and... cyclical?), vessel # (the count of vessels), ADFG # (individual vessel (fixed or random effect??)), Season aka year, 
--If I do not include jdate, simplifies things?
Response: nominal cpue

Q: should I do something besides a log-transformation?
```{r}

#cpue dist
ggplot(D7) + aes(x=CPUE_nom) + geom_density()
ggplot(D7) + aes(x=log(CPUE_nom+0.001)) + geom_density()
qqnorm(log(D7$CPUE_nom)) #that kind of looks bad. Maybe somehting other than log-trans?
qqnorm(D7$CPUE_nom)

ggplot(D7) + aes(x=CPUE_nom) + geom_density()+facet_wrap(~Analysis.Area)
ggplot(D7) + aes(x=log(CPUE_nom+0.001)) + geom_density()+facet_wrap(~Analysis.Area)

#cpue by year
ggplot(D7) + aes(x=factor(Season.Ref), y=CPUE_nom) + geom_point() #ther is a big outlier
ggplot(D7) + aes(x=factor(Season.Ref), y=CPUE_nom) + geom_boxplot(outliers = F)

ggplot(D7) + aes(x=factor(Season.Ref), y=CPUE_nom) + geom_boxplot(outliers = F) + facet_wrap(~Analysis.Area) #cpue varies... cyclically with time. Is season a fixed effect? I want to estimate it, so probs not. its def not linear.

#catch by year
ggplot(D7) + aes(x=factor(Season.Ref), y=total_weight) + geom_boxplot() #this is total weight per fish ticket.
ggplot(D7) + aes(x=factor(Season.Ref), y=max_pots_2) + geom_boxplot()
ggplot(D7) + aes(x=factor(Season.Ref), y=max_pots_2) + geom_boxplot(outliers=F)

ggplot(D7) + aes(x=factor(Season.Ref), y=total_weight) + geom_boxplot() + facet_wrap(~Analysis.Area) #max pots per fish ticket
ggplot(D7) + aes(x=factor(Season.Ref), y=max_pots_2) + geom_boxplot(outliers=F) + facet_wrap(~Analysis.Area)


#cpue and vessel count
ggplot(D7) + aes(x=factor(vessel_count_mgmt_u), y=log(CPUE_nom)) + geom_boxplot()
ggplot(D7) + aes(x=factor(vessel_count_mgmt_u), y=log(CPUE_nom)) + geom_boxplot(outliers=F)
ggplot(D7) + aes(x=factor(vessel_count_mgmt_u), y=log(CPUE_nom)) + geom_boxplot(outliers=F) + facet_wrap(~Analysis.Area)

#cpue and district


#cpue and jdate



#correlation plot
library(corrplot)
D7_interest_cor <- D7 %>% select(CPUE_nom, jdate, vessel_count_mgmt_u)
cor_prep <- cor(D7_interest_cor)
corrplot(cor_prep)
?pairs

```


Analysis and model testing: district 7 GAM with random effects
```{r}
#null model
m0 <- gam(log(CPUE_nom + 0.001) ~ factor(Season.Ref) ,data= D7 ) #how to add temporal autocorrealtion to this tho?

#intermediate models


#global model
m_glob_r <- gam(log(CPUE_nom + 0.001) ~ factor(Season.Ref) + factor(ADFG.Number)+ s(vessel_count_mgmt_u, k=4) + s(jdate, k=4) + s(factor(Analysis.Area), bs="re", k=4), data=D7 ) #remove jdate? #crap should ADFG.Number be a\random effect too?
##do I need to make this a factor somewhere else?
test_df <- D7
test_df$Analysis.Area <- factor(test_df$Analysis.Area)
m_glob_r <- gam(log(CPUE_nom + 0.001) ~ factor(Season.Ref) + factor(ADFG.Number)+ s(vessel_count_mgmt_u, k=4) + s(jdate, k=4) + s(Analysis.Area, bs="re", k=4), data=test_df) #ok that worked. I jsut cant make it a factor within my equation. Maybe I should adjust my df code accordingly.
summary(m_glob_r)

##how to alter the random effect to have a random global slope? See that paper.
m_glob_nr
  
  
#first decide if ranef or not, via global model?


#TEST BELWO
test <- gam(log(CPUE_nom+0.001) ~ vessel_count_mgmt_u + Season.Ref + ADFG.Number + s(jdate, k=4), data=(D7)) #that worked
test2 <- gam(log(CPUE_nom+0.001) ~ vessel_count_mgmt_u + Season.Ref + ADFG.Number + s(jdate, k=4) + s(factor(Analysis.Area), bs="re"), data=(D7)) #that did not work. Soemthign abotu my random effect not working.
```

Analysis and model testing: individual analysis areas
```{r}
#Bradfield canal

#Lower ernest sound

#upper ernest sound

#Zimovia strait


```

Comparison graphs
```{r}

```

Results graphs
```{r}

```

