---
title: "District 2"
author: "Alex Reich"
date: "2024-05-14"
output: html_document
---
Set the wd for this rmd
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/agreich/OneDrive - State of Alaska/Desktop/Shrimp/Shrimp tickets (pot shrimp)/SE_pot_shrimp_CPUE')
```


Load libraries
```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(mgcv)
```


Load the data and select for D1
```{r}
wrangled_shrimp <- read.csv("Data/wrangled shrimp focus years.csv")
D2 <- wrangled_shrimp %>% filter(Management_unit == "District 2") %>%
  mutate(Analysis.Area = factor(Analysis.Area),
         ADFG.Number = factor(ADFG.Number),
         Season.Ref = factor(Season.Ref)
         )#make things factors for the gams

unique(D2$Analysis.Area)

```


Exploratory data analysis
```{r}

```


Add area weights for model selection
- see pg 113 of campbell 2015
- Nobs(total # of observations)/Nstrata(which is Nyear*Nanalysisarea) * 1/# observations in the analysis area within the year and region (??)
- or Eobs/Nstrata * 1/e_ijk THIS ONE.
- how to weight??
- I do not get this.
```{r}

```


Model selection
- check if interaction between year and area
- campbell 2015 says use gamma with log link instead of gaussian with log link. Try both.
- area and yera should be fixefs, but test vessel ID as ranef (it will probs also be a fixef)
- consider adding the year:area interaction as a ranef (which would be the slope, not int I think)
-- run that overnight
```{r}

```

Model reiduals
```{r}

```


Calculate standarized cpue with predict() and newdata
- use mode of vessel for the vessel effect, but check the mode of the vessel in all of the areas
- see tyler email about setting up newdata
```{r}

```

Graphs
```{r}

```

Total std CPUE in region 1 per year (later, once all calc'ed)
sum(area_k*CPUE_ik) where i is each year and k is each area

Average b
