---
title: "District 7"
author: "Alex Reich"
date: "2024-04-24"
output: html_document
---

Set RMD directory
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/agreich/OneDrive - State of Alaska/Desktop/Shrimp/Shrimp tickets (pot shrimp)/SE_pot_shrimp_CPUE')
```


Key points: 6/4/24:
- QC my wrangle please, and give me clean data next time b/c wrangling takes forever
- how I dealt with missing data- last object moved forward
- model: global and selected
- analysis by mgmt area -> predict the results by analysis area (using the mode vessel to each area), average to mgmt unit, average mgmt units to get Region-wide cpue- using the mgmt unit level!!
--more efficient than 67 or whatever models per analysis area.
-model structure: global and selected model
- from the wish list: what are you talking about? Examples please.
-How will these results be used? -how will this cpue standardization be used
- decision making level? at the distrist/mgmt u or analysis area
-same page about me passing off this code, check in in ~5 years
- survey dates?


Anyway, let's get started. This document will look at management unit District 7 for pot (spot) shrimp standardization. It will load in the (pre-wrangled) data, wrangle further for district 1, and then some EDA (focusing on variables of interest and correlation). I will then make graphs of the standardized CPUE (by district and) and whatever else. THEN, I will run individual models by analysis areas, ideally with an automated function, and compare the graphs of those automated functions to the cpue of the ranef model.This will help me decide if models should be by MANAGEMENT UNIT (larger) or by ANALYSIS AREA (smaller). 


Load libraries
```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(mgcv)
```


Load the data and select for D7
```{r}
wrangled_shrimp <- read.csv("Data/wrangled shrimp focus years.csv")
D7 <- wrangled_shrimp %>% filter(Management_unit == "District 7") %>%
  mutate(Analysis.Area = factor(Analysis.Area),
         ADFG.Number = factor(ADFG.Number),
         Season.Ref = factor(Season.Ref)
         )#make things factors for the gams

unique(D7$Analysis.Area)

```


District 7 contains analysis areas:
  Bradfield Canal
  Lower Ernest Sound
  Upper Ernest Sound
  Zimovia Strait
  
  
  
## Exploratory Data Analysis for D7
Variables of interest:
Random effect: Analysis.Area
Fixed effects: jdate (smoothed and... cyclical?), vessel # (the count of vessels), ADFG # (individual vessel (fixed or random effect??)), Season aka year, 
--If I do not include jdate, simplifies things?
Response: nominal cpue

Q: should I do something besides a log-transformation?
THIN OUT
```{r}

#cpue dist
ggplot(D7) + aes(x=CPUE_nom) + geom_density()
ggplot(D7) + aes(x=log(CPUE_nom+0.001)) + geom_density()
qqnorm(log(D7$CPUE_nom)) #that kind of looks bad. Maybe somehting other than log-trans?
qqnorm(D7$CPUE_nom)

#ggplot(D7) + aes(x=CPUE_nom) + geom_density()+facet_wrap(~Analysis.Area)
ggplot(D7) + aes(x=log(CPUE_nom+0.001)) + geom_density()+facet_wrap(~Analysis.Area)

#cpue by year
#ggplot(D7) + aes(x=factor(Season.Ref), y=CPUE_nom) + geom_point() #ther is a big outlier
ggplot(D7) + aes(x=factor(Season.Ref), y=log(CPUE_nom)) + geom_boxplot(outliers = F)

ggplot(D7) + aes(x=factor(Season.Ref), y=log(CPUE_nom)) + geom_boxplot(outliers = F) +  facet_wrap(~Analysis.Area) #cpue varies... cyclically with time. Is season a fixed effect? I want to estimate it, so probs not. its def not linear.

#catch by year
ggplot(D7) + aes(x=factor(Season.Ref), y=total_weight) + geom_boxplot() #this is total weight per fish ticket.
ggplot(D7) + aes(x=factor(Season.Ref), y=max_pots_2) + geom_boxplot()
ggplot(D7) + aes(x=factor(Season.Ref), y=max_pots_2) + geom_boxplot(outliers=F)

ggplot(D7) + aes(x=factor(Season.Ref), y=total_weight) + geom_boxplot() + facet_wrap(~Analysis.Area) #max pots per fish ticket
ggplot(D7) + aes(x=factor(Season.Ref), y=max_pots_2) + geom_boxplot(outliers=F) + facet_wrap(~Analysis.Area)


#cpue and vessel count
ggplot(D7) + aes(x=factor(vessel_count_mgmt_u), y=log(CPUE_nom)) + geom_boxplot()
ggplot(D7) + aes(x=factor(vessel_count_mgmt_u), y=log(CPUE_nom)) + geom_boxplot(outliers=F)
ggplot(D7) + aes(x=factor(vessel_count_mgmt_u), y=log(CPUE_nom)) + geom_boxplot(outliers=F) + facet_wrap(~Analysis.Area)

#cpue and district


#cpue and jdate



#correlation plot
#library(corrplot)
#D7_interest_cor <- D7 %>% select(CPUE_nom, jdate, vessel_count_mgmt_u)
#cor_prep <- cor(D7_interest_cor)
#corrplot(cor_prep)
#?pairs


#check model residuals, see if we need ar1 correlation in the model
```


Drop the outlier, perhaps
```{r}
D7 <- D7 %>% filter(CPUE_nom < 100) #gonna assume that is an outlier
```


Remove ADFG vessel from consideration (add this to shrimp prep function later)
```{r}
D7 <- D7 %>% filter(ADFG.Number!= 99999)
unique(D7$ADFG.Number)
```



Model selection 5/15/24- non-imputed data
```{r}

#ok can breifly test ranefs
M_ranef <-  lmer(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + Season.Ref:Analysis.Area + Analysis.Area + (1|ADFG.Number), data=D7, REML=T)


M_noranef <-  lm(log(CPUE_nom+0.001) ~ Season.Ref +vessel_count_aa + ADFG.Number + Season.Ref:Analysis.Area + Analysis.Area, data=D7)
AIC(M_ranef, M_noranef) #no ranefs

summary(M_ranef)



#global with ranef and interactions

M2 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + ADFG.Number +Analysis.Area + Season.Ref:Analysis.Area, data=D7)

M3 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + ADFG.Number + Analysis.Area + Season.Ref:Analysis.Area, data=D7)

M4 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + ADFG.Number +Analysis.Area, data=D7)

M5 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + Analysis.Area + Season.Ref:Analysis.Area, data=D7)

M_null <- lm(log(CPUE_nom+0.001) ~ Season.Ref+ Analysis.Area, data=D7)

AIC(M2, M3, M4, M5, M_null) #M3 wins
#says that vessel count does not matter...


summary(M3)

```





Imputation method: LOCF- last obervation carried forwards 
```{r}
D7$logCPUE <- log(D7$CPUE_nom) #I do not need the +0.001 because everything is positive/ not 0


temp2 <- expand.grid(Season.Ref = unique(D7$Season.Ref),
                     Analysis.Area = unique(D7$Analysis.Area)
)
D7_sub <- D7 %>% select(Season.Ref, Analysis.Area, ADFG.Number, logCPUE)
D7_sub2 <- full_join(D7_sub, temp2)
#make missing values for ADFG.Number the mode
which.max(table(D7_sub2$ADFG.Number)) #52131

mode_boat <- which.max(table(D7$ADFG.Number))

mode_val <- names(mode_boat)
  

D7_sub3 <- D7_sub2 %>%
  mutate(ADFG.Number = replace_na(ADFG.Number, mode_val)) #give the missing value the mode fishing vessel

############################################
#filling in the missing year value with the mean of the last year
##############################################
  # Step 1: Calculate the mean logCPUE for each season
season_means <- D7_sub3 %>%
  group_by(Season.Ref, Analysis.Area) %>%
  summarise(mean_logCPUE = mean(logCPUE, na.rm = TRUE)) %>%
  arrange(Season.Ref, Analysis.Area)
  
  # Step 2: Join the season means back to the original dat
  D7_filled_2 <- D7_sub3 %>%
  left_join(season_means)
  
  # Step 3: Fill the NA values with the mean from the previous season
D7_filled_2 <- D7_filled_2 %>%
  arrange(Season.Ref) %>%
  group_by(Analysis.Area) %>%
  mutate(logCPUE = if_else(is.na(logCPUE),
                           dplyr::lag(mean_logCPUE),
                           logCPUE)) %>%
  ungroup()

# Drop the mean_logCPUE column as it's no longer needed
D7_filled_2 <- D7_filled_2 %>%
  select(-mean_logCPUE)

ggplot(D7_filled_2) + aes(x=Season.Ref, y=logCPUE) + geom_point() + facet_wrap(~Analysis.Area)

  #ok, that looks good,
  
  
```


Model selection after imputation
```{r}
#D7_filled_2 is my dataframe
#import areas of .... stat areas
## from an R standpoint, it will be messy and full of NA's
areas_raw <- read.csv("Data/Southeast_Shrimp_StatArea_SqMiles.csv")
names(areas_raw)
areas_D7 <- areas_raw %>% #from the messy dataset
  select(DISTRICT_CODE, STAT_AREA_NAME, STAT_AREA, Area_SqMiles) %>% #choose the columns I care about
  filter(DISTRICT_CODE == 107) #%>% #I only care about District 7 right now
  #group_by(S)

#View(areas_D7)

U_ern_sound <- areas_D7 %>% filter(STAT_AREA == 10720) %>%
  summarise(area=sum(Area_SqMiles))
L_ern_sound <- areas_D7 %>% filter(STAT_AREA == 10710) %>%
  summarise(area=sum(Area_SqMiles))
Zim_st <- areas_D7 %>% filter(STAT_AREA == 10730|STAT_AREA == 10735) %>%
  summarise(area=sum(Area_SqMiles))
Brad_Can <- areas_D7 %>% filter(STAT_AREA == 10740|STAT_AREA == 10745) %>%
  summarise(area=sum(Area_SqMiles))

df_temp <- data.frame(Analysis.Area= c("Upper Ernest Sound", "Lower Ernest Sound", "Bradfield Canal", "Zimovia Strait"),
                      area_sqmi = c(U_ern_sound$area, L_ern_sound$area, Zim_st$area, Brad_Can$area))

#ugh that was tedious
##at least I have my areas now

D7_filled_3 <- left_join(D7_filled_2, df_temp) #joined by analysis areas, now I have areas that I can use to weigh

#generate a model
M3_2 <- lm(logCPUE~ Season.Ref + ADFG.Number + Analysis.Area + Season.Ref:Analysis.Area, data=D7_filled_3) 
summary(M3_2)
M_alt1 <- lm(logCPUE~ Season.Ref + ADFG.Number + Analysis.Area, data=D7_filled_3)
M_alt2 <- lm(logCPUE~ Season.Ref + Analysis.Area + Season.Ref:Analysis.Area, data=D7_filled_3) 

AIC(M3_2, M_alt1, M_alt2) #M3_2 wins

plot(M3_2)

```




Predict -global vessel mode - D7 only
```{r}
table(D7$ADFG.Number)


std_dat<- expand.grid(Season.Ref = unique(D7_filled_3$Season.Ref), #predict with every boat as the mode for D7(52131)
                              Analysis.Area = unique(D7_filled_3$Analysis.Area), 
                              ADFG.Number = factor(52131) 
                      )

pred_logcpue <- predict(M3_2, std_dat, type = "response", se = TRUE) 


#get the unlogged cpue and put back in DF
ln_mu = pred_logcpue$fit
ln_sigma = pred_logcpue$se.fit

std_dat %>% 
   mutate(
          ln_mu=ln_mu,
          mu = exp(ln_mu + (ln_sigma^2)/2),
          sigma = sqrt((exp(ln_sigma^2) - 1) * exp(2*ln_mu + ln_sigma^2)), #make sure is right
          upper= mu + 2*sigma,
          lower = mu - 2*sigma
   ) -> pred_dat_biascorrected #gives the predicted dataset for using the same mode boat for all areas



```

From tyler code:
# f_bias_correct () ----
# bias correction from log space
# args: ln_mu = mean of log(x)
#       ln_sigma = standard error of log(x)
f_bias_correct <- function(ln_mu, ln_sigma) {
  
  mu = exp(ln_mu + (ln_sigma^2)/2)
  sigma = sqrt((exp(ln_sigma^2) - 1) * exp(2*ln_mu + ln_sigma^2))
  
  return(c(mu = mu, sigma = sigma))
}


Standardized (predicted)  CPUE
(from internet: Note I predict the log version as exp(predict(m) + sigma^2/2) because $E exp(X) = exp(mu + sigma^2)$ where $X \sim N(0, sigma^2)$. (This changes little here though.))
See: https://stats.stackexchange.com/questions/115571/back-transforming-regression-results-when-modeling-logy





Analysis and model testing: individual analysis areas- for D7 comparison only
```{r}
#separate datasets (use the datasets where missing values already filled)
Ernest_U <- D7 %>% filter(Analysis.Area == "Upper Ernest Sound")
Ernest_L <- D7_filled_3 %>% filter(Analysis.Area == "Lower Ernest Sound")
Bradfield <- D7_filled_3 %>% filter(Analysis.Area == "Bradfield Canal")
Zim <- D7_filled_3 %>% filter(Analysis.Area == "Zimovia Strait")

#model testing per dataset (use step...?)
ErnestU_mod1 <- lm(logCPUE ~ Season.Ref + ADFG.Number + vessel_count_aa, Ernest_U)
step_EU <- step(ErnestU_mod1) #gets rid of vessel effect

ErnestL_mod1 <- lm(logCPUE ~ Season.Ref + ADFG.Number, Ernest_L)
step_EL <- step(ErnestL_mod1)

Brad_mod1 <- lm(logCPUE ~ Season.Ref + ADFG.Number, Bradfield)
step_brad<-step(Brad_mod1)

Zim_mod1 <- lm(logCPUE ~ Season.Ref + ADFG.Number, Zim)
step_zim<-step(Zim_mod1)

#predictions
##u enrest
D7_pred <- function(DATA, MODE_VESSEL, MODEL){
std_dat<- expand.grid(Season.Ref = unique(DATA$Season.Ref),
                              Analysis.Area = unique(DATA$Analysis.Area), 
                              ADFG.Number = factor(MODE_VESSEL) #maybe pick mode vessel for this area. This is the mode vessel for D7 overall
                      )
pred_logcpue <- predict(MODEL, std_dat, type = "response", se = TRUE) 
ln_mu = pred_logcpue$fit
ln_sigma = pred_logcpue$se.fit

std_dat %>% 
   mutate(
          ln_mu=ln_mu,
          mu = exp(ln_mu + (ln_sigma^2)/2),
          sigma = sqrt((exp(ln_sigma^2) - 1) * exp(2*ln_mu + ln_sigma^2)), #make sure is right
          upper= mu + 2*sigma,
          lower = mu - 2*sigma
   ) -> pred_dat_biascorrected

return(pred_dat_biascorrected)

}

#max boat
mode_boat <- which.max(table(D7$ADFG.Number))

mode_val <- names(mode_boat)

EL_mode_boat <- names(which.max(table(Ernest_L$ADFG.Number)))
brad_mode_boat <- names(which.max(table(Bradfield$ADFG.Number)))
zim_mode_boat<- names(which.max(table(Zim$ADFG.Number)))
#function in action

EU_pred <- D7_pred(DATA=Ernest_U, MODE_VESSEL = 52131, MODEL = step_EU)
EL_pred <- D7_pred(DATA=Ernest_L, MODE_VESSEL = 71757, MODEL = step_EL) #why do you not work
Brad_pred <- D7_pred(DATA=Bradfield, MODE_VESSEL = 14980, MODEL = step_brad)
Zim_pred <- D7_pred(DATA=Zim, MODE_VESSEL = 70552, MODEL = step_brad)

comb_D7 <- rbind(EU_pred, Brad_pred, Zim_pred) #EL_pred is being finicky
#comb_D7 <- rbind(comb_D7, Zim_pred)


#oh wait. I can just take the coeffs per year... right? since there are no interactions
#maybe not via the method I chose tho


```

Comparison graphs
```{r}
#D7 master
ggplot(pred_dat_biascorrected) + aes(x=Season.Ref, y=mu) +geom_errorbar(aes(ymin=lower, ymax=upper))+ geom_point() + geom_line(aes(group=Analysis.Area)) + facet_wrap(~Analysis.Area)

#overall D7 cpue #ok need to work on... sum the cpue's by area to get the total cpue's per year.. what about the error here tho?
#ggplot(pred_dat_biascorrected) + aes(x=Season.Ref, y=mean(mu)) +geom_errorbar(aes(ymin=lower, ymax=upper))+ geom_point() + geom_line()

#individual separations
ggplot(comb_D7) + aes(x=Season.Ref, y=mu) +
  geom_errorbar(aes(ymin=lower, ymax=upper))+ 
  geom_point() + geom_line(aes(group=Analysis.Area)) + facet_wrap(~Analysis.Area) 

##########################3
#just look at upper ernest
ern_u_from_large_D7 <- pred_dat_biascorrected %>% filter(Analysis.Area =="Upper Ernest Sound")
ggplot(ern_u_from_large_D7) + aes(x=Season.Ref, y=mu) +geom_errorbar(aes(ymin=lower, ymax=upper))+ geom_point() + geom_line(aes(group=Analysis.Area))# + facet_wrap(~Analysis.Area)

ggplot(EU_pred) + aes(x=Season.Ref, y=mu) +
  geom_errorbar(aes(ymin=lower, ymax=upper))+ 
  geom_point() + geom_line(aes(group=Analysis.Area))# + facet_wrap(~Analysis.Area)

##pretty similar from one to the otehr
###let's look at bradfield and Zim now:
brad_u_from_large_D7 <- pred_dat_biascorrected %>% filter(Analysis.Area =="Bradfield Canal")
ggplot(brad_u_from_large_D7) + aes(x=Season.Ref, y=mu) +geom_errorbar(aes(ymin=lower, ymax=upper))+ geom_point() + geom_line(aes(group=Analysis.Area))# + facet_wrap(~Analysis.Area)

ggplot(Brad_pred) + aes(x=Season.Ref, y=mu) +
  geom_errorbar(aes(ymin=lower, ymax=upper))+ 
  geom_point() + geom_line(aes(group=Analysis.Area))

zim_u_from_large_D7 <- pred_dat_biascorrected %>% filter(Analysis.Area =="Zimovia Strait") #are we sure this is not bradfield? check filtering
ggplot(zim_u_from_large_D7) + aes(x=Season.Ref, y=mu) +geom_errorbar(aes(ymin=lower, ymax=upper))+ geom_point() + geom_line(aes(group=Analysis.Area))# + facet_wrap(~Analysis.Area)

ggplot(Zim_pred) + aes(x=Season.Ref, y=mu) +
  geom_errorbar(aes(ymin=lower, ymax=upper))+ 
  geom_point() + geom_line(aes(group=Analysis.Area))

#and this tells a diff story...
##max makes decisons by analysis areas?
##Can include stat area if I go lower mode
##what if I modeled mode vessel for each specific area? Can I do that?

#OH! I can predict invidudually for each area using the gloval model, and choose the mode vessel per area. That should help!!!
```

Well, I don't like how different the graphs are between the full D7 model and separate by area models. I can fix by using the vessel modes separated by area, I think. Like this:
```{r}



##predict function
D7_pred_2 <- function(DATA, ANALYSIS_AREA, MODE_VESSEL, MODEL){
std_dat<- expand.grid(Season.Ref = unique(DATA$Season.Ref),
                              Analysis.Area = ANALYSIS_AREA, 
                              ADFG.Number = factor(MODE_VESSEL) #maybe pick mode vessel for this area. This is the mode vessel for D7 overall
                      )
pred_logcpue <- predict(MODEL, std_dat, type = "response", se = TRUE) 
ln_mu = pred_logcpue$fit
ln_sigma = pred_logcpue$se.fit

std_dat %>% 
   mutate(
          ln_mu=ln_mu,
          mu = exp(ln_mu + (ln_sigma^2)/2),
          sigma = sqrt((exp(ln_sigma^2) - 1) * exp(2*ln_mu + ln_sigma^2)), #make sure is right
          upper= mu + 2*sigma,
          lower = mu - 2*sigma
   ) -> pred_dat_biascorrected

return(pred_dat_biascorrected)

}

ernU_pred <- D7_pred_2(MODEL = M3_2, ANALYSIS_AREA = "Upper Ernest Sound", MODE_VESSEL = 52131, DATA=D7_filled_3)
ernL_pred <- D7_pred_2(MODEL = M3_2, ANALYSIS_AREA = "Lower Ernest Sound", MODE_VESSEL = EL_mode_boat, DATA=D7_filled_3)
brad_pred <- D7_pred_2(MODEL = M3_2, ANALYSIS_AREA = "Bradfield Canal", MODE_VESSEL = brad_mode_boat, DATA=D7_filled_3)
zim_pred <- D7_pred_2(MODEL = M3_2, ANALYSIS_AREA = "Zimovia Strait", MODE_VESSEL = zim_mode_boat, DATA=D7_filled_3)

D7_pred_comb <- rbind(ernU_pred, ernL_pred, brad_pred, zim_pred)

D7_pred_comb <- left_join(D7_pred_comb, df_temp)

```


Try graphing THAT
```{r}

ggplot(D7_pred_comb) + aes(x=Season.Ref, y=mu) + geom_errorbar(aes(ymin=lower, ymax=upper)) + geom_point() + geom_line(aes(group="Analysis.Area"))+
  facet_wrap(~Analysis.Area)

#how does this compare to my other method?

#the combined method where I use the overall vessel mode:
ggplot(pred_dat_biascorrected) + aes(x=Season.Ref, y=mu) +geom_errorbar(aes(ymin=lower, ymax=upper))+ geom_point() + geom_line(aes(group=Analysis.Area)) + facet_wrap(~Analysis.Area)

#the combined plot when I separate things:
ggplot(comb_D7) + aes(x=Season.Ref, y=mu) +
  geom_errorbar(aes(ymin=lower, ymax=upper))+ 
  geom_point() + geom_line(aes(group=Analysis.Area)) + facet_wrap(~Analysis.Area) 


##close look at zimovia
a<- D7_pred_comb %>% filter(Analysis.Area =="Zimovia Strait") %>%
  ggplot() + aes(x=Season.Ref, y=mu) + geom_errorbar(aes(ymin=lower, ymax=upper)) + geom_point()+geom_line(aes(group="Analysis.Area"))

b <- pred_dat_biascorrected %>% filter(Analysis.Area =="Zimovia Strait") %>%
  ggplot() + aes(x=Season.Ref, y=mu) + geom_errorbar(aes(ymin=lower, ymax=upper)) + geom_point()+geom_line(aes(group="Analysis.Area")) 

c <- comb_D7 %>% filter(Analysis.Area =="Zimovia Strait") %>%
  ggplot() + aes(x=Season.Ref, y=mu) + geom_errorbar(aes(ymin=lower, ymax=upper)) + geom_point() + geom_line(aes(group="Analysis.Area"))

library(patchwork)
a+b+c

#close look at bradfield
D7_pred_comb %>% filter(Analysis.Area =="Bradfield Canal") %>%
  ggplot() + aes(x=Season.Ref, y=mu) + geom_errorbar(aes(ymin=lower, ymax=upper)) + geom_point()+geom_line(aes(group="Analysis.Area"))

pred_dat_biascorrected %>% filter(Analysis.Area =="Bradfield Canal") %>%
  ggplot() + aes(x=Season.Ref, y=mu) + geom_errorbar(aes(ymin=lower, ymax=upper)) + geom_point()+geom_line(aes(group="Analysis.Area")) 

comb_D7 %>% filter(Analysis.Area =="Bradfield Canal") %>%
  ggplot() + aes(x=Season.Ref, y=mu) + geom_errorbar(aes(ymin=lower, ymax=upper)) + geom_point() + geom_line(aes(group="Analysis.Area"))

#close look at U ernest

```

Ok, based on these graphs I've decided area-wide analysis is fine, and I'd like to use the mode of each vessel for analysis. Now make the graphs for D7 and do the other stuff.



Results graphs
D7_pred_comb separates to: ernU_pred, ernL_pred, brad_pred, zim_pred
```{r}

#make teh function (move to functions page later)
make.plots <- function(DATA_PRED, AREA_NAME, DATA_RAW){ #data pred is the
 (E1<-ggplot(data=DATA_PRED) + aes(x=Season.Ref, y=mu) + 
   geom_point(size=3)+
   geom_errorbar( aes(ymin=lower, ymax=upper)) +
   geom_line(aes(group=1))+ #that works+
   #geom_hline(aes(yintercept=mean(std_dat_ranef_lim$bt_cpue)), linetype="dashed")+
   theme_cowplot()+
   labs(y="Standardized CPUE (lbs/pots)", x=NULL, title = AREA_NAME )+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))#+
   #  scale_y_continuous(breaks = c(1,2,3,4,5,6), expand=c(0,0))+
    # ylim(1.8, 5)
   )

 raw_totals <- DATA_RAW %>%
   group_by(Season.Ref, Analysis.Area) %>%
   dplyr::summarize(Harvest = sum(total_weight),
                    Effort = sum(max_pots_2))
raw_totals %>% filter(Analysis.Area == AREA_NAME) %>%
ggplot() + aes(x=Season.Ref) +
   geom_point(aes(y=Harvest), shape=0, size=3)+ #harvest graph
   geom_line(aes(y=Harvest, group=1))+
   geom_point(aes(y=Effort), shape=15, size=3) +
   geom_line(aes(y=Effort, group=1), linetype="dashed") +
   theme_cowplot()+
   labs(x="Season", y="Harvest (lbs) or Effort (pots)")+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
E2 <- last_plot()

(Plot <- E1/E2)
return(Plot)
}

#upper ERnest
Upper_ernest_plot <- make.plots(DATA_PRED = ernU_pred, AREA_NAME = "Upper Ernest Sound", DATA_RAW=D7)
Lower_ernest_plot <- make.plots(DATA_PRED = ernL_pred, AREA_NAME = "Lower Ernest Sound", DATA_RAW=D7)
Brad_plot <- make.plots(DATA_PRED = brad_pred, AREA_NAME = "Bradfield Canal", DATA_RAW=D7)
Zim_plot <- make.plots(DATA_PRED = zim_pred, AREA_NAME = "Zimovia Strait", DATA_RAW=D7)

Upper_ernest_plot
Lower_ernest_plot
Brad_plot
Zim_plot

```



D7 district-wide graphs
Average of analysis area cpue's, weighted by area
Weighted and unweighted graphs
```{r}

#D7 df's
##combine values
##Looking at p 111 of campbell 2015
D7_results_area_weighted <- D7_pred_comb %>%
  group_by(Season.Ref) %>% #I want the sum by year
  summarise(CPUE_avg = weighted.mean(x=mu, w=area_sqmi),
            CPUE_sd = sqrt(weighted.mean(x=sigma^2, w=area_sqmi))
  )%>%
   mutate(upper = CPUE_avg + 2*CPUE_sd,
         lower = CPUE_avg - 2*CPUE_sd)
  # area = max(area_sqmi)) %>% #ok this part is fucked. Fix it. I think it calcs the biomass anyway so mayve dont worry
  #mutate(index= CPUE_sum*area) #average? or sum? Hoyle et al says take the wighted sum
#Calc se!!!!!

#library(car)$
#car::deltaMethod(M3_2, mean())
#library(msm)
#deltamethod(g=mean(D7_pred_comb$mu), mean=mean(D7_pred_comb$mu), cov = summary(M3_2)$sigma^2)


#6/4/24
##get mean of sd
D7_pred_comb$sigma #my standard deviations
var_D7<-D7_pred_comb$sigma^2 #my variances
sqrt(mean(var_D7))
mean(D7_pred_comb$sigma) #these are different, good to know
#only works for equal sample sizes, fuck. nWait, in predict I DO have equal sample sizes





#model:M3_2
summary(M3_2)


D7_results_not_weighted <- D7_pred_comb %>%
  group_by(Season.Ref) %>% #I want the sum by year
  summarise(CPUE_avg = mean(mu), #average the cpue's
            #CPUE_sum = sum(mu), #I do not think I need this
            CPUE_sd = sqrt(mean(sigma^2))) %>%
  mutate(upper = CPUE_avg + 2*CPUE_sd,
         lower = CPUE_avg - 2*CPUE_sd)#get the sd's by taking the mean of the variances, then converting back to sd.



#total D7 plot
 raw_totals_D7 <- D7 %>%
   group_by(Season.Ref) %>%
   dplyr::summarize(Harvest = sum(total_weight),
                    Effort = sum(max_pots_2))
raw_D7_plot <-raw_totals_D7 %>%# filter(Analysis.Area == AREA_NAME) %>%
 ggplot() + aes(x=Season.Ref) +
   geom_point(aes(y=Harvest), shape=0, size=3)+ #harvest graph
   geom_line(aes(y=Harvest, group=1))+
   geom_point(aes(y=Effort), shape=15, size=3) +
   geom_line(aes(y=Effort, group=1), linetype="dashed") +
   theme_cowplot()+
   labs(x="Season", y="Harvest (lbs) or Effort (pots)")+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
#D7 graphs unweighted (soo... weighted (or biased, depending on your perspective) by catch)
#D7_unweighed <- make.plots(DATA_PRED = D7_pred_comb,AREA_NAME = NULL, DATA_RAW = D7 ) #no area name, will it work? #NOPE. graph by hand
D7_no_weight <- ggplot(D7_results_not_weighted)  + aes(x=Season.Ref, y=CPUE_avg) + 
   geom_point(size=3)+
   geom_errorbar( aes(ymin=lower, ymax=upper)) +
   geom_line(aes(group=1))+ #that works+
   #geom_hline(aes(yintercept=mean(std_dat_ranef_lim$bt_cpue)), linetype="dashed")+
   theme_cowplot()+
   labs(y="Standardized CPUE (lbs/pots)", x=NULL, title = "District 7" )+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   # scale_y_continuous(breaks = c(1,2,3,4,5,6), expand=c(0,0))+
    ylim(0, 8)

D7_plot <- D7_no_weight/raw_D7_plot

#D7 graphs weighted by area
D7_yes_weight <- ggplot(D7_results_area_weighted)  + aes(x=Season.Ref, y=CPUE_avg) + 
   geom_point(size=3)+
   geom_errorbar( aes(ymin=lower, ymax=upper)) +
   geom_line(aes(group=1))+ #that works+
   #geom_hline(aes(yintercept=mean(std_dat_ranef_lim$bt_cpue)), linetype="dashed")+
   theme_cowplot()+
   labs(y="Standardized CPUE (lbs/pots)", x=NULL, title = "District 7" )+
   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   # scale_y_continuous(breaks = c(1,2,3,4,5,6), expand=c(0,0))+
    ylim(0, 8)

D7_results_area_weighted
D7_results_not_weighted

D7_plot
(D7_plot_weight <- D7_yes_weight/raw_D7_plot)
```




This is where shrimp survey data is located.
https://www.adfg.alaska.gov/analytics/saw.dll?Answers&path=%2Fshared%2FCommercial%20Fisheries%2FRegion%20I%2FInvertebrates%2FSurvey%2FShrimp%2FSpecimens%20with%20Pot%20Data

Survey catches are also reported in fish ticket data, recorded as Fishery: Test Fish, and Permit Holder: ADFG Sitka. These survey catches should be removed from calculations in CPUE and total reportable harvest.- as of 6/5/24 not included in wrangle function from shrimp_functions.R- add it in.



