---
title: "District 1"
author: "Alex Reich"
date: "2024-05-08"
output: html_document
---

Set the wd for this rmd
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/agreich/OneDrive - State of Alaska/Desktop/Shrimp/Shrimp tickets (pot shrimp)/SE_pot_shrimp_CPUE')
```


Load libraries
```{r}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(mgcv)
```


Load the data and select for D1
```{r}
wrangled_shrimp <- read.csv("Data/wrangled shrimp focus years.csv")
D1 <- wrangled_shrimp %>% filter(Management_unit == "District 1") %>%
  mutate(Analysis.Area = factor(Analysis.Area),
         ADFG.Number = factor(ADFG.Number),
         Season.Ref = factor(Season.Ref)
         )#make things factors for the gams

unique(D1$Analysis.Area)

```


Exploratory plots
```{r}
#cpue dist
ggplot(D1) + aes(x=CPUE_nom) + geom_density()
ggplot(D1) + aes(x=log(CPUE_nom+0.001)) + geom_density()

qqnorm(log(D1$CPUE_nom))


#variables vs year
ggplot(D1) + aes(x=factor(Season.Ref), y=CPUE_nom) + geom_boxplot()
ggplot(D1) + aes(x=factor(Season.Ref), y=CPUE_nom) + geom_boxplot(outliers=F)
ggplot(D1) + aes(x=factor(Season.Ref), y=log(CPUE_nom+ 0.001)) + geom_boxplot()
ggplot(D1) + aes(x=factor(Season.Ref), y=log(CPUE_nom+ 0.001)) + geom_boxplot() + facet_wrap(~Analysis.Area)
ggplot(D1) + aes(x=factor(Season.Ref), y=log(CPUE_nom+ 0.001)) + geom_boxplot(outliers=F) + facet_wrap(~Analysis.Area)
ggplot(D1) + aes(x=factor(Season.Ref), y=log(CPUE_nom+ 0.001)) + geom_boxplot() + facet_wrap(~Analysis.Area)
ggplot(D1) + aes(x=factor(Season.Ref), y=log(CPUE_nom+ 0.001), color=Analysis.Area) + geom_boxplot() #+ facet_wrap(~Analysis.Area)
#does cpue slope vary with analysis area? Totally does:
ggplot(D1) + aes(x=Batch.Year, y=log(CPUE_nom+ 0.001), color=Analysis.Area) + geom_smooth(se=F) #+ geom_point()
ggplot(D1) + aes(x=Batch.Year, y=log(CPUE_nom+ 0.001), color=Analysis.Area) + geom_smooth(method="lm", se=F)
#does cpue slope vary with ADFG.ID?
ggplot(D1) + aes(x=Batch.Year, y=log(CPUE_nom+ 0.001), color=ADFG.Number) + geom_smooth(method="lm", se=F) + theme(legend.position="none") #hello absolute chaos

ggplot(D1) + aes(x=factor(vessel_count_mgmt_u), y=log(CPUE_nom)) + geom_boxplot(outliers=F) + facet_wrap(~Analysis.Area)
ggplot(D1) + aes(x=vessel_count_mgmt_u, y=log(CPUE_nom)) + geom_smooth() + facet_wrap(~Analysis.Area)
ggplot(D1) + aes(x=vessel_count_mgmt_u, y=log(CPUE_nom)) + geom_smooth() #+ facet_wrap(~Analysis.Area)
ggplot(D1) + aes(x=vessel_count_mgmt_u, y=log(CPUE_nom)) + geom_smooth(method="lm") + geom_point()

#so medium amound of vessel might be less cpue, but I think this is correlated with fishing conditions in the year

#adfg number and cpue
ggplot(D1) + aes(y= log(CPUE_nom+0.001), x=ADFG.Number) + geom_boxplot()

#does the trend in log cpue vary by analysis area?


```

Model selection DO NOT RUN (RUN OVERNIGHT. To tell me if I need random effects)
```{r}
#throw a gam in there
mod_g_new <- gam(log(CPUE_nom+0.001) ~ Season.Ref + s(vessel_count_aa, k=4) + s(Analysis.Area, bs="re") + s(ADFG.Number, bs="re"), method = "ML", data=D1)
mod_g_2 <- gam(log(CPUE_nom+0.001) ~ Season.Ref + s(Analysis.Area, bs="re") + s(ADFG.Number, bs="re"), method = "ML", data=D1)

AIC(mod_g_new, mod_g_2) #smoothed term not needed, so do not use a gam...

#determine random effect structure
M_1 <-  lmer(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + Season.Ref:vessel_count_aa + (1|ADFG.Number) + (1|Analysis.Area), data=D1, REML=F) #global with ranef and interactions
M_2 <- lmer(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + Season.Ref:vessel_count_aa + ADFG.Number + (1|Analysis.Area), data=D1, REML=F)
M_3 <- lmer(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + Season.Ref:vessel_count_aa + (1|ADFG.Number) + Analysis.Area, data=D1, REML=F)
M_4 <- lmer(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + Season.Ref:vessel_count_aa + (1 + Season.Ref|ADFG.Number) + (1+ Season.Ref|Analysis.Area), data=D1, REML=F) #full model, this may take a really long time to run and not actually work out.
M_5 <- lmer(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + Season.Ref:vessel_count_aa + (1|ADFG.Number) + (1+ Season.Ref|Analysis.Area), data=D1, REML=F)
M_6 <- lmer(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + Season.Ref:vessel_count_aa + (1 + Season.Ref|ADFG.Number) + (1|Analysis.Area), data=D1, REML=F)
M_7 <-  lmer(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + Season.Ref:vessel_count_aa + ADFG.Number + (1 + Season.Ref|Analysis.Area), data=D1, REML=F) 
M_8 <-  lmer(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + Season.Ref:vessel_count_aa + (1+Season.Ref|ADFG.Number) + Analysis.Area, data=D1, REML=F) 
M_9 <-  lmer(log(CPUE_nom+0.001) ~ Season.Ref * vessel_count_aa + Analysis.Area + (1|ADFG.Number), data=D1, REML=F) #no, cant account for the int effects
M_10 <- lm(log(CPUE_nom+0.001) ~ Season.Ref * vessel_count_aa * ADFG.Number*Analysis.Area, data=D1) #fixef only with ALL the interaction effects #M15, same as
M10_adjust <- lm(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + ADFG.Number +Analysis.Area + Season.Ref:Analysis.Area, data=D1)

summary(M10_adjust)

AIC(M_1, M_2, M_3, M_4, M_5, M_6, M_7, M_8, M_9, M_10)
#for the record, based on theory I think that analysis area should be a random effect and MAYBE ADFG.ID as well
##thing might get too complex once we start doing random slope, if anything has a random slope it should be Analysis.Area/

#determine fixed effect structure HMM WHAT DID CURRY DO AGAIN?!?!?!

AIC(M10_adjust, M_1, M_2, M_3) #the not random effects wins


############


```


Could use bam and bam syntax to cut down run time
https://m-clark.github.io/posts/2019-10-20-big-mixed-models/
```{r}
mod_b_1 <- bam(log(CPUE_nom+0.001) ~ Season.Ref + s(Analysis.Area, bs="re") + s(ADFG.Number, bs="re"), method = "ML", data=D1)
mod_b_2 <- bam(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + s(Analysis.Area, bs="re") + s(ADFG.Number, bs="re")+
                 s(Season.Ref, Analysis.Area, bs="re") + s(Season.Ref, ADFG.Number, bs="re"), method = "ML", data=D1, discrete=T, nthreads = 16)
```


Ok let's ignore ranefs and look at fixefs
```{r}
M10_adjust <- lm(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + ADFG.Number +Analysis.Area + Season.Ref:Analysis.Area, data=D1)

#M11 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + vessel_count_aa + ADFG.Number +Analysis.Area + Season.Ref:Analysis.Area+ Season.Ref:ADFG.Number, data=D1) #takes forever

M12 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + ADFG.Number +Analysis.Area + Season.Ref:Analysis.Area, data=D1)

M13 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + ADFG.Number +Analysis.Area, data=D1)

M14 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + Analysis.Area + Season.Ref:Analysis.Area, data=D1)

AIC(M10_adjust, M12, M13, M14) #go with M12, there's interaction effects

summary(M12)
```

Run these models overnight- DO NOT RUN, 
too many singularities
```{r}
M15 <- lm(log(CPUE_nom+0.001) ~ Season.Ref * ADFG.Number * Analysis.Area, data=D1)
#M16 <- lm(log(CPUE_nom+0.001) ~ Season.Ref + ADFG.Number + Analysis.Area + ADFG.Number:Analysis.Area, data=D1
#stats::step(M12) #test
stats::step(M15)
summary(M15)

AIC(M12, M15)

```




Model residuals graph
```{r}
plot(M12)
plot(M15)
```



Predicted results column
```{r}
predictions <- predict(M12, newdata = D1, type = "response", interval ="confidence", se.fit=T) #can I just use D1 as the new data or is that cheating
#see line 808 of phils's Jig_BRF_CPUE code. he went with averaged dataset

D1_new <- D1
D1_new$std_CPUE <- exp(predictions$fit[,1])-0.001
D1_new$lower_CI <- exp(predictions$fit[,2])-0.001
D1_new$upper_CI <- exp(predictions$fit[,3]) -0.001
D1_new$se <- exp(predictions$se.fit)-0.001

head(D1_new) #that my new dataset, with the standardized CPUE (backt tranformed predicted values)


#predict results: cpue by year for D1
coef(M12) #"extract your year coefs, those will be the annual index. Need to log-tranform them back?
str(coef(M12))
exp(coef(M12) + 0.001)
#"Scale the coeffs to center on 1
##soo then
M12_adj <- lm(log(CPUE_nom+0.001) ~ 0 + Season.Ref + ADFG.Number +Analysis.Area + Season.Ref:Analysis.Area, data=D1) #remove the intercept
###um, should I do it this way? Tyler had things a different way in teh scallop cpue stuff, mayve check that out? Something about changing the default option in lm
summary(M12_adj)
exp(coef(M12_adj) + 0.001)



```



Graphs

Graph of standardized cpue


other graph


Combine

